\chapter{Introduction}

\section{Motivation and Background}

hello

\section{Problem Description}

Describe the problem we want to solve

\section{Problem Formulation}

Mathematical formulation of problem

\section{Overall Organization of Work}

ok

\section{Delimitations}

important!

Be very descriptive to make sure I cant be deducted for not looking into X.

\section{Structure of the Report}

\renewcommand{\descriptionlabel}[1]{\normalfont #1}
\begin{description}[itemsep=2pt, topsep=0pt, parsep=0pt]
    \item[Chapter 1] presents the motivation for the work, defines the problem, and describes the organisation of the project.

    \item[Chapter 2] introduces the theoretical background for point cloud processing and related geometric operations and provides the essential knowledge to follow the literature study and the developed algorithms.

    \item[Chapter 3] describes the literature study and explains how published work informs the choices in this project.

    \item[Chapter 4] gives an overview of the full software pipeline and explains the main processing stages.

    \item[Chapter 5] describes the detailed methods used to detect compartments, segment structural elements, and build the scene graph.

    \item[Chapter 6] presents the datasets used to develop and test the system.

    \item[Chapter 7] presents the results from each stage of the pipeline and identifies the weaknesses and strengths of the algorithms.

    \item[Chapter 8] concludes the report by summarizing the contributions and achievements.
\end{description}

