\section{Optimization Methods}
\label{sec:theory_optmet}

In practice, it is often hard to solve a nonlinear residual problem directly. Instead, the problem can be reformulated as a nonlinear least squares optimization problem with the scalar objective function
\begin{equation}
f(x) = \frac{1}{2} \| r(x) \|^2 \in \mathbb{R}.
\label{eq:least_squares_objective}
\end{equation}
Solving the minimization problem,
\begin{equation}
\label{eq:min_problem}
    x^\ast = \underset{x}{\text{argmin}} \ f(x) = \underset{x}{\text{argmin}} \ \frac{1}{2} \| r(x) \|^2,
\end{equation}
satisfies $r(x^*) = 0$ if an exact solution exists. If no exact solution exists, or if the system is overdetermined, the formulation yields a solution that minimizes the residual norm.\\

Solving \cref{eq:min_problem} is very hard to solve in general \cite{madsen2004}, since there might exist local minima that are larger than a global minima. \cite{nielsen1999}

A group of methods used to solve the optimization problem operates by iteratively evaluating the objective function at the current iterate $x_k$ and determining a step $s_k$ that decreases the cost function. The next iterate is computed as
\begin{equation}
\label{eq:step_update_rule}
    x_{k+1} = x_k + s_k,
\end{equation}
with the goal of choosing $s_k$ so that
\begin{equation*}
    f(x_{k+1}) < f(x_k).
\end{equation*}
There are two main frameworks for controlling how the step $s_k$ is chosen, called \textit{Line Search} and \textit{Trust Region} methods. This section will introduce these frameworks along with common algorithms used to solve the minimization problem.

\subsection{Line Search Methods}
\begin{equation}
    s_k = \alpha_k p_k
\end{equation}
\begin{itemize}
    \item Armijo Backtracking\\
    \item Wolfe Conditions
\end{itemize}

\subsection{Gradient Descent}
Gradient descent is a simple method that is based on moving in the direction of the negative gradient of the objective function. 
Using the chain rule, the gradient of $f$ is given by
\begin{equation}
\label{eq:gradient_f}
    \nabla f(x) = J(x)^\top r(x),
\end{equation}
where $J(x)$ denotes the Jacobian of the residual function.

At the current iterate $x_k$, the search direction is chosen as the direction of steepest descent:
\begin{equation}
\label{eq:steepest_descent}
    p_k = - \nabla f(x_k) = - J(x_k)^\top r(x_k).
\end{equation}

Gradient descent is simple to implement and requires only first-order derivative information through the Jacobian. It is computationally inexpensive per iteration and applicable even when second-order derivatives are unavailable. However, the method may converge slowly, especially for ill-conditioned problems or narrow curved valleys in the objective function. Its performance is sensitive to scaling and may require many iterations to achieve high accuracy.

\subsection{Newton's Method}

Newton's method incorporates second-order information of the objective function. 
The gradient of $f$ is given by \cref{eq:gradient_f} and the Hessian is given by
\begin{equation}
\label{eq:hessian}
    \nabla^2 f(x) = J(x)^\top J(x) + \sum_{i=1}^m r_i(x)\nabla^2 r_i(x).
\end{equation}

At iteration $k$, the Newton search direction is obtained by solving the linear system
\begin{equation*}
    \nabla^2 f(x_k) p_k = - \nabla f(x_k).
\end{equation*}

By incorporating curvature information, Newton's method typically achieves quadratic convergence in a neighborhood of a solution, provided that the Hessian is nonsingular. This represents a significant improvement over gradient descent in terms of local convergence speed.

However, the method requires computation and factorization of the Hessian matrix, which can be expensive for large-scale problems. Moreover, if the Hessian is indefinite or poorly conditioned, the method may fail to produce a descent direction without additional safeguards.

\subsection{Gauss--Newton Method}

The Gauss--Newton method is tailored specifically to nonlinear least squares problems. 
It is based on an approximation of \cref{eq:hessian},
\begin{equation*}
    \nabla^2 f(x) \simeq J(x)^\top J(x),
\end{equation*}
which neglects the second-order term involving $\nabla^2 r_i(x)$.

The search direction is therefore computed by solving
\begin{equation}
\label{eq:gauss_newton}
    J(x_k)^\top J(x_k) p_k = - J(x_k)^\top r(x_k).
\end{equation}

This approximation avoids explicit computation of second derivatives and reduces computational cost compared to full Newton's method. When the residual is small near the solution, the neglected term becomes insignificant, and the method exhibits fast local convergence.

The Gauss--Newton method may perform poorly when the residual is large or when $J(x_k)^\top J(x_k)$ is rank deficient. In such cases, the approximation of the Hessian can be inaccurate, and additional regularization techniques may be required to ensure robustness.

\subsection{Trust Region Methods}
\label{sec:trust_region_methods}

Trust region methods determine the step $s_k$\footnote{The step $s_k$ is the same as the search direction $p_k$ with $\alpha=1$.} by minimizing a local model of the objective function within a restricted neighborhood around the current iterate. From \cite{madsen2004}, a typical model is the second order taylor expansion, using the Gauss--Newton approximation of the Hessian, 
\begin{equation}
\label{eq:mk}
f(x_k+s) \simeq m_k(s) = f(x_k) + r(x_k)^\top J(x_k) \, s + \frac{1}{2} s^\top J(x_k)^\top J(x_k) \, s
\end{equation}

In the trust region method, it is assumed that there exists a radius $\Delta$ around $x_k$, such that the model $m_k(s)$ is sufficiently accurate within the region. The step is then computed by solving the constrained minimization problem,

\begin{equation}
\label{eq:trust_region_subproblem}
    s_k = \underset{s}{\text{argmin}} \ m_k(s) \quad \text{subject to} \quad \|s\| \leq \Delta_k.
\end{equation}

In practice, this is solved through solving a damped problem, where the step is computed by solving the unconstrained minimization problem,
\begin{equation}
\label{eq:damped_problem}
    s_k = \underset{s}{\text{argmin}} \ \left(m_k(s) + \frac{1}{2} \ \lambda \ s^\top s\right),
\end{equation}
where $\lambda \geq 0$ is the damping parameter, and $\lambda \ s^\top s$ penalizes large steps. 

For a given $\Delta$ or $\lambda$, if no sufficient step $s_k$ is found such that $f(x_k+s_k)<f(x_k)$ (assuming $x_k$ is not a solution), the parameters ($\Delta,\lambda$) must be tuned such that a sufficient step is found. Since $m_k(s)$ is assumed to be a sufficient approximation for $f(x_k+s)$ when $s$ is sufficiently small, the trust region should be decreased ($\lambda$ increased) for the next iteration. When a sufficient step is found, it indicates that $m_k(s)$ is a good approximation, and the trust region can be increased ($\lambda$ decreased), aiming for faster convergence.

\subsection{Levenberg--Marquardt Method} \label{sec:levenberg_marquardt}

The Levenberg--Marquardt (LM) method is a damped Gauss--Newton method designed for solving nonlinear least--squares problems. It can be interpreted as a trust region method where the step is computed from a regularized Gauss--Newton system. The damping parameter, $\lambda$, controls the transition between the Gauss--Newton and steepest descent method.

At iteration $k$, the LM step is obtained by solving the linear system
\begin{equation} \label{eq:lm_system}
    \left( J(x_k)^\top J(x_k) + \lambda I \right) s_k
    =
    - J(x_k)^\top r(x_k),
\end{equation}
As $\lambda \to 0$, the system approaches the Gauss--Newton equations, \cref{eq:gauss_newton}. For large $\lambda$, the step approaches,
\begin{equation*}
    s_k \simeq -\frac{1}{\lambda}J(x_k)^\top r(x_k),
\end{equation*}
which is the steepest descent direciton, \cref{eq:steepest_descent}, with a small step lenght, $\alpha=\lambda^{-1}$.

The damping parameter must be initialized before the iterations begin. \cite{nielsen1999} states that the initial value should relate to the magnitude of the eigenvalues, and proposes
\begin{equation} \label{eq:init_lambda}
\lambda_0 = \tau \max \left( \operatorname{diag}(J(x_0)^\top J(x_0)) \right),
\end{equation}
where $\tau > 0$ is a small constant.

To evaluate the quality of a proposed step $s_k$, the LM method uses the gain ratio, defined as the ratio between the actual reduction in the objective function and the reduction predicted by the quadratic model,
\begin{equation}
\label{eq:gain_ratio}
\varrho =
\frac{f(x_k) - f(x_k + s_k)}
     {m_k(0) - m_k(s_k)}.
\end{equation}

The denominator can be computed from \cref{eq:mk},
\begin{align}
    m_k(0)-m_k(s_k) &= -s_k^\top J(x_k)^\top r(x_k) \,  - \frac{1}{2} s_k^\top J(x_k)^\top J(x_k) \, s_k \notag \\
    &= -\frac{1}{2} s_k^\top(2 J(x_k)^\top r(x_k) + (J(x_k)^\top J(x_k) + \lambda I-\lambda I)s_k) \notag \\
    &= \frac{1}{2} s_k^\top (\lambda s_k-J(x_k)^\top r(x_k)) \label{eq:mk_solve}.
\end{align}

Combining \cref{eq:gain_ratio,eq:mk_solve} yields a computationally easier expression for the gain ratio,
\begin{equation}
    \label{eq:gain_ratio_solve}
    \varrho = \frac{\|r(x_k)\|^2 - \|r(x_k+s_k)\|^2}
    {s_k^\top (\lambda s_k-J(x_k)^\top r(x_k))}.
\end{equation}
Note that $s_k^\top s_k$ and $-s_k^\top J(x_k)^\top r(x_k)$ are positive, such that the denominator is always positive.

If $\varrho > 0$, the step is accepted and the iterate is updated,
\[
x_{k+1} = x_k + s_k.
\]
The damping parameter is then decreased according to
\begin{equation}
\label{eq:dec_lambda}
\lambda := 
\lambda \max \left\{ \frac{1}{3},\, 1 - (2\varrho - 1)^3 \right\}.
\end{equation}

If $\varrho \le 0$, the step is rejected, and the damping parameter is increased by
\begin{equation}
\label{eq:inc_lambda}
\lambda := 2^{i+1}\lambda,
\end{equation}
where $i\in \{0,1,\dots,i_{\max}\}$ is the iteration counter until the step is accepted. This update strategy is demonstrated in \cite{nielsen1999}, and is shown to be a sufficient choice.\\

Intuitively, when $\varrho$ is positive, the quadratic model $m_k(s)$ provides a good approximation of the objective function, and the damping parameter is reduced so that the method behaves more like Gauss--Newton. When $\varrho$ is negative, the model is considered unreliable, and the damping parameter is increased, producing a smaller and more conservative step.


\subsubsection{Computing the step with SVD factorization}
\cref{eq:lm_system} can be solved by SVD factorization. Using,
\begin{equation*}
    J(x_k) = U_k \Sigma_k V_k^\top,
\end{equation*}
and performing some algebraic manipulation, the step can be computed as
\begin{equation}
\label{eq:sk_lm_solve}
    s_k = -V_k \, \text{diag}(d_i) \, U_k^\top r(x_k), 
\end{equation}
Where $d_i=\dfrac{\sigma_i}{\sigma_i^2 + \lambda}$, and $\sigma_i$ are the singular values from $\Sigma_k$.\\

\cref{eq:sk_lm_solve} computes the LM step without computing $J^\top J$, which can be numerically unstable when $J$ is ill-conditioned.


