\chapter{Discussion on Findings}
\label{chap:discussion}

This chapter discusses the thesis's main findings, the practical limitations of the proposed method, and directions for further work.

\section{Main Findings}

The results demonstrate that steady-state inlet and outlet measurements can contain enough information to identify more than two simultaneous leaks under idealized conditions. This is a stronger result than has been demonstrated by the closest steady-state methods reviewed in \autoref{chap:literature}. At the same time, the results show that the information available from endpoint measurements is limited and that this limitation becomes apparent before the number of unknowns becomes very large.\\

The main reason is that each additional leak introduces two new parameters, while the effect of each parameter is only observed indirectly through the outlet head and flow. As the number of leaks increases, different changes in leak locations and leak coefficients can produce increasingly similar changes in the residual. This is reflected in the Jacobian, where the columns become more aligned as $n$ increases. The solver behavior should therefore be interpreted as a consequence of the inverse problem itself, not only as a property of the implemented optimization method.\\

This also explains why the choice of operating points is important. Measurements taken under similar hydraulic conditions add little independent information, while measurements with a larger pressure-drop range improve conditioning. However, the benefit of adding more measurements diminishes as $m$ increases. Additional steady-state measurements are therefore most useful when they represent sufficiently different operating points, not merely because they increase the number of equations.\\

The results for four leaks show the difference between exact identification and useful localization. The solver does not always recover the true configuration within the strict tolerance $\varepsilon=0.001$, and several test cases produce multiple solutions. Nevertheless, many of the incorrect solutions are close to the true leak configuration, as shown in \autoref{sec:err_mag}. This indicates that the endpoint measurements may still contain useful localization information even when they are insufficient to uniquely recover all leak parameters with high precision.\\

For five leaks, the true solution is rarely recovered, and the false detection rate becomes high. Together with the conditioning analysis, this suggests that four leaks is close to the useful numerical limit of the proposed approach under the tested assumptions. This is not an absolute theoretical limit, since some leak configurations are easier to identify than others, but it indicates that identifying many simultaneous leaks from boundary measurements alone is unlikely to be reliable without additional information.\\

The results for an unknown number of leaks support a more practical interpretation of the method. Solving the problem with $\hat{n} = 2$ or $\hat{n} = 3$ does not recover the full leak configuration when additional leaks are present, but it often produces an estimated location close to that of one true leak. The method may therefore be useful for reducing the search distance in an iterative procedure in which a leak is located and repaired before new measurements are collected.

\section{Practical Limitations}

The strongest practical limitation is the assumption that the pipe parameters are known. In particular, the resistance coefficient \gls{theta} has a direct influence on the predicted relation between inlet and outlet measurements. In practice, \gls{theta} must often be estimated from pipe documentation, friction models, or calibration data. For old pipes without historical monitoring, leak-free calibration data may not be available, and the effective resistance may differ from the assumed value.\\

The sensitivity analysis shows that this limitation is important. Small errors in \gls{theta} can be compensated by large changes in the estimated leak parameters, which means that an inaccurate model can produce plausible but incorrect leak configurations. The residual may be useful for searching over possible values of \gls{theta} in simple cases, but the reliability of this approach is uncertain when measurement noise and other model errors are present at the same time.\\

Measurement errors create a similar difficulty. Random noise can be reduced by averaging steady-state samples, but bias errors and deviations from exact steady-state conditions may remain. Since the method relies on small differences between measured and predicted endpoint values, the solver can interpret such errors as changes in leak locations or leak sizes. This is especially problematic for configurations that are already poorly conditioned.\\

The practical usefulness of the method, therefore, depends on more than solver convergence. It requires sufficiently accurate pipe parameters, reliable endpoint measurements, and operating points that produce enough variation in the hydraulic response. The current method is therefore better viewed as an idealized identification study and a possible aid to localization, rather than a standalone field method.

\section{Future Work}

The most important next step is to improve robustness to modeling errors and measurement noise. One direction is to investigate whether pipe parameters such as \gls{theta} and \gls{ell} can be estimated together with the leak parameters. This would reduce the dependence on exact prior knowledge, but it would also increase the number of unknowns and make the inverse problem more difficult. It is therefore important to determine whether the endpoint measurements contain enough information for such a formulation.\\

Experimental or field data are also needed to assess practical performance. Such data would show whether the steady-state model is sufficiently accurate, how large the measurement errors are, and whether the required variation between operating points can be obtained in practice. This would provide a more realistic test of the method than simulated measurements alone.\\

The results from \autoref{sec:n_not_known} also suggest that future work should investigate strategies based on several assumed values of $\hat n$. Combining the results from multiple solver configurations may provide a more robust way to reduce the search distance when the true number of leaks is unknown.\\

Finally, extending the method to small pipe networks with junctions would make the formulation more relevant to real water distribution systems. Such an extension would likely introduce additional identifiability challenges and should therefore be studied together with the conditioning and robustness issues identified in this thesis.
